### 2018-03-02 code/heterogeneity.rmd
Standardized RRR8

---
title: "Heterogeneity in pregregistered multi-lab studies"
author: "Anton Ohlsson Collentine"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r load data}
library(readr)

dat <- read_csv("../data/collated_summary_data.csv")
```

```{r metafor}
library(metafor)
library(dplyr)
library(purrr)

#this function is for transforming RRR8 which only has raw mean difference and SE to a standardized version
transform_SE <- function(ES, SE, n1, n2){ #Based on meta-analysis lecture 02
  sdpooled <- sqrt(SE^2 / (1 / n1 + 1/n2))
  d <- ES / sdpooled
  g <- (1 - 3 / (4*(n1 + n2) - 9))*d
  v <- 1 / n1 + 1/ n2 + g^2 / (2*(n1 + n2))
  data.frame(g = g, v = v)
}

est_heterogen_smd_raw <- function(x){
  if(any(x[, "outcomes1_2"] == "mean _ SD")){ #without the 'any' we will get warnings because we apply 'if' to a vector

    fit <- rma(measure = "SMD", m1i = outcome_t1, m2i = outcome_c1, sd1i = outcome_t2, sd2i = outcome_c2, #gives Hedge's g as outcome
               n1i = ntreatment, n2i = ncontrol,test = "knha", data = x)

    fit2 <- rma(measure = "MD", m1i = outcome_t1, m2i = outcome_c1, sd1i = outcome_t2, sd2i = outcome_c2, #gives raw outcome
               n1i = ntreatment, n2i = ncontrol,test = "knha", data = x)

  } else if(any(x[, "outcomes1_2"] == "mean _ SE")){

    standardized <- transform_SE(x$effect_size, x$outcome_c2, x$ntreatment, x$ncontrol)

    fit <- rma(yi = g, vi = v, test = "knha", data = standardized) #standardized effect estimate

    fit2 <-  rma(yi = effect_size, sei = outcome_c2, test = "knha",  data = x) #raw effect estimate

  } else if(any(x[, "effect_type"] == "r")){

    fit <- fit2 <- rma(measure = "COR", ri = effect_size, ni = Ntotal, test = "knha", data = x)

    I2_1 <- confint(fit2)$random[3, ]
    I2_2 <- rep(NA, 3)
    #In this case, stop function and return dataframe below
    return(data.frame(s_I2 = I2_1[1], s_ci.lb = I2_1[2], s_ci.ub = I2_1[3], r_I2 = I2_2[1], r_ci.lb = I2_2[2], r_ci.ub = I2_2[3])) #output is for smd, r, and 1 raw difference

  } else{ #for all two-group count effects

    fit <- rma(measure = "OR2D", ai = outcome_t1, bi = outcome_t2, ci = outcome_c1, di = outcome_c2, #standardized
               n1i = ntreatment, n2i = ncontrol, test = "knha", data = x)

    fit2 <- rma(measure = "RD", ai = outcome_t1, bi = outcome_t2, ci = outcome_c1, di = outcome_c2, #unstandardized
               n1i = ntreatment, n2i = ncontrol, test = "knha", data = x)
  }

  I2_1 <- confint(fit)$random[3, ] #Gives us the I2 estimate and its confidence interval for standardized values
  I2_2 <- confint(fit2)$random[3, ] #Gives us the I2 estimate and CI for raw values

  data.frame(s_I2 = I2_1[1], s_ci.lb = I2_1[2], s_ci.ub = I2_1[3], r_I2 = I2_2[1], r_ci.lb = I2_2[2], r_ci.ub = I2_2[3]) #output is for smd, r, and 1 raw difference
}

res <- dat %>% 
  filter(!rs == "RRR5") %>% #Remove RRR5 for now because have not been able to extract sufficient data for it
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map_dfr(est_heterogen_smd_raw, .id = "effect") %>% #apply function, rbind results into dataframe ('dfr'), 
  mutate(dif = s_I2 - r_I2) #Gives us the difference between the two

#Add rs to the data-frame
het <- dat %>% 
  select(rs, effect) %>% 
  group_by(rs, effect) %>% 
  summarize() %>% 
  left_join(res) %>% 
  arrange(rs, desc(s_I2)) #sort by rs and I2 for standardized effect

```

```{r table}
library(kableExtra)
het %>% knitr::kable("latex", booktabs = T, digits = 1, col.names = c("RS", "Effect", "I2(%)", "ci.lb", "ci.ub", "I2(%)", "ci.lb", "ci.ub", "Dif")) %>% 
  kable_styling() %>% 
  add_header_above(c(" " = 2, "Standardized" = 3, "Unstandardized" = 3, " " = 1))

```
**Note**: I am uncertain if we can speak of standardized/unstandardized in regards to 2x2 tables? For now these correspond to OR/Risk difference for the following five effects {Allowed vs forbidden, gain vs. loss framing, norm of reciprocity, low vs. high category scale, verbal overshadowing}. NA in unstandardized columns comes from correlations which don't seem possible to calculate the unstandardized version for with the available information. RRR5 I have not yet been able to extract a measure of variance for, hence the NAs.