### 2018-07-08 writing/appendixA-heterogeneity.rmd
Added results for simulation sensitivity check of effect size

---
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.asp = 0.618)
```

```{r load packages, data and source helper functions}
if(!require(readr)){install.packages("readr")}
if(!require(dplyr)){install.packages("dplyr")}
if(!require(ggplot2)){install.packages("ggplot2")}
if(!require(purrr)){install.packages("purrr")}
if(!require(metafor)){install.packages("metafor")}

library(readr) #To load data
library(dplyr) #For data transformation
library(ggplot2) #Plot data
library(purrr) #For data iteration
library(metafor) #To run meta-analyses

dat <- read_csv("../data/collated_summary_data.csv")

source("../code/helper_functions_tables_figures.r") #Load functions to prep data for figures
```

```{r results}
##Zero effect size results
dens <- readRDS("../data/power_simulation_results.RDS") 
names(dens) <-  sort(unique(dat$effect))

#Summary of results. These are incorporated into the main table in the paper, see tables.rmd
zero <- dens %>% 
  bind_rows(.id = "effect") %>% 
  group_by(effect, tau_index) %>% 
  summarize(power = mean(Qp <= 0.05)) %>% #Estimate power/type 1 error for each tau level and effect
  ungroup() %>% 
  tidyr::spread(key = tau_index, value = power) %>% #prep for table
  rename(Zero = '1', Small = '2', Medium = '3', Large = '4', Effect = effect)

##Medium effect size results
dens2 <- readRDS("../data/appendixA_power_simulation_results.RDS") 
names(dens2) <-  sort(unique(dat$effect))

#Summary of results.
medium <- dens2 %>% 
  bind_rows(.id = "effect") %>% 
  group_by(effect, tau_index) %>% 
  summarize(power = mean(Qp <= 0.05)) %>% #Estimate power/type 1 error for each tau level and effect
  ungroup() %>% 
  tidyr::spread(key = tau_index, value = power) %>% #prep for table
  rename(Zero = '1', Small = '2', Medium = '3', Large = '4') %>% 
  select(-effect)

##combine for table

comparison <- cbind(zero, medium)
```

Appendix A - Sensitivity analysis of effect size for simulations

In our main results we simulated data to estimate type I error and power using an average effect size of zero. As explained in the methods section this is unlikely to matter as our interest lies in heterogeneity. However, as a sensitivity analysis we in this Appendix again estimate type I error and power, but generated with an average 'medium' effect size as defined by Cohen (1988).

Estimated type I error and power for zero/small/medium/large heterogeneity when simulated with a true effect size of zero vs. medium are shown for each effect in Table A1. In both cases type I error is approximately nominal, as compared to the expected 5% error rate. With a an effect size of zero, average power was `r 100*round(mean(zero$Small), digits = 2)`%/`r 100*round(mean(zero$Medium), digits = 2)`%/`r 100*round(mean(zero$Large), digits = 2)`% for small/medium/large heterogeneity respectively. The corresponding power with an average effect size was `r 100*round(mean(zero$Small), digits = 2)`%/ `r 100*round(mean(zero$Medium), digits = 2)`%/`r 100*round(mean(zero$Large), digits = 2)`%. To conclude, changing true effect size in our simulations from zero to medium had little to no impact on estimated type I error and power.

Table A1.

_Type I error and power to detect four heterogeniety levels under two true effect sizes_

```{r}
knitr::kable(comparison, digits = 2)
```

**Note.** Column headers indicate the degree of simulated heterogeneity where Zero = simulated type 1 error, and the other headers represent simulated power under small/medium/large heterogeneity ($I^2$ = 25/50/75%) respectively. Medium effect size is as defined by Cohen (1988). Code to reproduce table: osf.io/bsjhu/