### 2018-11-14 writing/appendixA-heterogeneity.rmd
Re-added appendices since one of them includes a table

---
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.asp = 0.618)
```

```{r load packages, data and source helper functions}
if(!require(readr)){install.packages("readr")}
if(!require(dplyr)){install.packages("dplyr")}
if(!require(ggplot2)){install.packages("ggplot2")}
if(!require(purrr)){install.packages("purrr")}
if(!require(metafor)){install.packages("metafor")}

library(readr) #To load data
library(dplyr) #For data transformation
library(ggplot2) #Plot data
library(purrr) #For data iteration
library(metafor) #To run meta-analyses

dat <- read_csv("../data/collated_summary_data.csv")

source("../code/helper_functions_tables_figures.r") #Load functions to prep data for figures
```

```{r results}
##Zero effect size results
dens <- readRDS("../data/power_simulation_results.RDS") 
names(dens) <-  sort(unique(dat$effect))

#Summary of results. These are incorporated into the main table in the paper, see tables.rmd
zero <- dens %>% 
  bind_rows(.id = "effect") %>% 
  group_by(effect, tau_index) %>% 
  summarize(power = mean(Qp <= 0.05)) %>% #Estimate power/type 1 error for each tau level and effect
  ungroup() %>% 
  tidyr::spread(key = tau_index, value = power) %>% #prep for table
  rename(Zero = '1', Small = '2', Medium = '3', Large = '4', Effect = effect)

##Medium effect size results
dens2 <- readRDS("../data/appendixA_power_simulation_results.RDS") 
names(dens2) <-  sort(unique(dat$effect))

#Summary of results.
medium <- dens2 %>% 
  bind_rows(.id = "effect") %>% 
  group_by(effect, tau_index) %>% 
  summarize(power = mean(Qp <= 0.05)) %>% #Estimate power/type 1 error for each tau level and effect
  ungroup() %>% 
  tidyr::spread(key = tau_index, value = power) %>% #prep for table
  rename(Zero = '1', Small = '2', Medium = '3', Large = '4') %>% 
  select(-effect)

##combine for table

comparison <- cbind(zero, medium)
```

Appendix A - Sensitivity analysis assuming a medium true effect size rather than zero true effect size

In our main results we simulated data to estimate type I error and power using an average true effect size equal to zero. As explained in the methods section this is unlikely to matter as our interest lies in heterogeneity. However, as a sensitivity analysis, in this appendix we also estimated type I error and power, but generated with an average 'medium' true effect size as defined by Cohen (1988).

Estimated type I error and power for zero/small/medium/large heterogeneity when simulated with a true effect size of zero vs. medium are shown for each effect in Table A1. In both cases type I error is approximately nominal, as compared to the expected 5% error rate. With a true effect size of zero, average power was `r 100*round(mean(zero$Small), digits = 2)`%/`r 100*round(mean(zero$Medium), digits = 2)`%/`r 100*round(mean(zero$Large), digits = 2)`% for small/medium/large heterogeneity, respectively. The corresponding power with a true medium effect size was also `r 100*round(mean(zero$Small), digits = 2)`%/ `r 100*round(mean(zero$Medium), digits = 2)`%/`r 100*round(mean(zero$Large), digits = 2)`%, respectively. To conclude, changing true effect size in our simulations from zero to medium had little to no impact on estimated type I error and power.

Table A1.

_Type I error and power to detect four heterogeneity levels under two true effect sizes_

```{r}
knitr::kable(comparison, digits = 2)
```

**Note.** Column headers indicate the degree of simulated heterogeneity where Zero = simulated type 1 error, and the other headers represent simulated power under small/medium/large heterogeneity ($I^2$ = 25/50/75%) respectively. Medium effect size is as defined by Cohen (1988). Code to reproduce table: osf.io/bsjhu/