### 2019-05-09 code/troubleshooting_tau2.rmd
Exploring tau2 - I2 relationship after transformation

---
title: "Troubleshooting tau2"
author: "Anton Olsson Collentine"
date: "7 mei 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load packages, data and source functions}
if(!require(readr)){install.packages("readr")}
if(!require(metafor)){install.packages("metafor")}
if(!require(dplyr)){install.packages("dplyr")}
if(!require(purrr)){install.packages("purrr")}
if(!require(kableExtra)){install.packages("kableExtra")}

library(readr) #To load data
library(dplyr) #For data transformation
library(purrr) #For data iteration
library(metafor) #To run meta-analyses
library(kableExtra) #For creating good-looking pdf tables
library(ggplot2)
library(cowplot)

dat <- read_csv("../data/collated_summary_data.csv")

source("helper_functions_tables_figures.r") #Load functions to prep data for figures
```

TO DO:
* First figure: replicate panel 2 but with untruncated
* Hopefully the last 3 points will move up
* Also compare mu for truncated and untruncated versions
* Maybe see if can figure out contribution of each lab to overall effect size mu

```{r, cache = TRUE}
res <- dat %>% 
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map_dfr(est_heterogen_smd_raw, .id = "effect") #apply function, rbind results into dataframe ('dfr'), make sure purrr is up to date

het <- dat %>% 
  select(rp, Site, effect, effect_type) %>% 
  group_by(rp, effect) %>% 
  summarize(effect_type = unique(effect_type),
            k = n_distinct(Site)) %>% 
  left_join(res) %>% 
  mutate(effect_type = recode(effect_type, 
                              d = "SMD.",
                              'Raw mean difference' = "MD",
                              'Risk difference' = "RD")) %>% 
  ungroup()

a <- het %>% filter(rp == "ML1" & !effect %in% c('Allowed vs. forbidden', 'Gain vs. loss framing','Norm of reciprocity', 'Low vs. high category scales') & effect_type == "SMD.")

a_plot <- ggplot(a, aes(x = s_I2, y = sqrt(tau2), label = effect)) + geom_point() + ggtitle("ML1 SMD - untransformed effect sizes") + geom_text()

```

```{r prepping, cache = TRUE}

#Transform effect sizes to biserial correlations (pearson's correlations left as is)
#estimate tau2 for correlations
#'transform_MA' and 'summarizer' are sourced functions
res2 <- dat %>% 
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map(transform_MA) %>% #
  map_dfr(summarizer, .id = "effect")

#Add rp and k to the data-frame and prep for table
het <- dat %>% 
  select(rp, Site, effect, effect_type) %>% 
  group_by(rp, effect) %>% 
  summarize(effect_type = unique(effect_type),
            k = n_distinct(Site)) %>% 
  left_join(res2) %>% 
  mutate(effect_type = recode(effect_type, 
                              d = "SMD.",
                              'Raw mean difference' = "MD",
                              'Risk difference' = "RD")) %>% 
  ungroup()

b <- het %>% filter(rp == "ML1" & !effect %in% c('Allowed vs. forbidden', 'Gain vs. loss framing','Norm of reciprocity', 'Low vs. high category scales') & effect_type == "SMD.")

b_plot <- ggplot(b, aes(x = I2, y = sqrt(tau2), label = effect)) + geom_point() + ggtitle("ML1 biserial r - transformed SMD effect sizes") + geom_text()

```

```{r}
cowplot::plot_grid(a_plot, b_plot, ncol = 1)
```

**The four effects at the end are Anchoring 1,2,3,4**

```{r, }
het %>% filter(rp == "ML3") %>% ggplot(aes(x = I2, y = sqrt(tau2))) + geom_point() + facet_wrap(~effect_type) + ggtitle("ML3 (SMD transformed to biserial correlations)")
```

**The one non-increasing effect in the left plot above is 'Credentials interaction'. **
However, note that all the ML3 'correlations' are either transformed partial eta-squared of 2x2 tables in the first table (we can't fix this, would have to get into the raw data which seems liek too much work, I mentioned this in the revised draft)

**Check which effects get truncated:**

```{r, echo = TRUE, cache = TRUE}
dat %>% 
  split(.$effect) %>% 
  map_dfr(transform_MA, .id = "effect") %>% 
  filter(r > 1)
```

Basically the anchoring effects.

```{r}
het %>% ggplot(aes(x = I2, y = sqrt(tau2))) + geom_point() + ggtitle("All effects - tetrachoric, biserial & product-moment r")
```
The effect that is close to one of the anchoring effects (I2 ~75, tau ~.05) is Allowed vs. forbidden which also has r = .90.