### 2019-07-11 code/Supplement-biserial-correlations.Rmd
Converted troubleshooting into supplement

---
title: "Supplement - Biserial correlation transformations"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
Sys.setlocale("LC_TIME", "English" ) #to print date in YAML in English
```

```{r load packages, data and source functions}
if(!require(readr)){install.packages("readr")}
if(!require(metafor)){install.packages("metafor")}
if(!require(dplyr)){install.packages("dplyr")}
if(!require(purrr)){install.packages("purrr")}

library(readr) #To load data
library(dplyr) #For data transformation
library(purrr) #For data iteration
library(metafor) #To run meta-analyses
library(ggplot2) #plotting figures
library(cowplot) #combine plots

dat <- read_csv("../data/collated_summary_data.csv")

source("helper_functions_tables_figures.r") #Load functions to prep data for figures
```

```{r a-plot-prep, cache = TRUE}
res <- dat %>% 
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map_dfr(est_heterogen_smd_raw, .id = "effect") #apply function, rbind results into dataframe ('dfr'), make sure purrr is up to date

het <- dat %>% 
  select(rp, Site, effect, effect_type) %>% 
  group_by(rp, effect) %>% 
  summarize(effect_type = unique(effect_type),
            k = n_distinct(Site)) %>% 
  left_join(res) %>% 
  mutate(effect_type = recode(effect_type, 
                              d = "SMD.",
                              'Raw mean difference' = "MD",
                              'Risk difference' = "RD")) %>% 
  ungroup()

a <- het %>% filter(rp == "ML1" & !effect %in% c('Allowed vs. forbidden', 'Gain vs. loss framing','Norm of reciprocity', 'Low vs. high category scales') & effect_type == "SMD.")
```

```{r, cache = FALSE}
a_plot <- ggplot(a, aes(x = s_I2, y = sqrt(tau2))) + geom_point() + ggtitle("SMD (untransformed)") +
  scale_x_continuous(name = expression(paste(italic(I)^2, "Index"))) +
  scale_y_continuous(name = "Tau")

```

```{r trunc-biserial, cache = TRUE}
#Transform effect sizes to biserial correlations (pearson's correlations left as is)
#estimate tau2 for correlations
#'transform_MA' and 'summarizer' are sourced functions
res2 <- dat %>% 
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map(transform_MA) %>% #
  map_dfr(summarizer, .id = "effect")

#Add rp and k to the data-frame 
het <- dat %>% 
  select(rp, Site, effect, effect_type) %>% 
  group_by(rp, effect) %>% 
  summarize(effect_type = unique(effect_type),
            k = n_distinct(Site)) %>% 
  left_join(res2) %>% 
  mutate(effect_type = recode(effect_type, 
                              d = "SMD.",
                              'Raw mean difference' = "MD",
                              'Risk difference' = "RD")) %>% 
  ungroup()

b <- het %>% filter(rp == "ML1" & !effect %in% c('Allowed vs. forbidden', 'Gain vs. loss framing','Norm of reciprocity', 'Low vs. high category scales') & effect_type == "SMD.")
```

```{r b-plot-1, cache = FALSE}
b_plot_1 <- ggplot(b, aes(x = I2, y = sqrt(tau2))) + geom_point() + ggtitle("Biserial r (transformed SMD)")

```

```{r untruncated-biserial, cache = TRUE}
res3 <- dat %>%
  filter(rp == "ML1" & !effect %in% c('Allowed vs. forbidden', 'Gain vs. loss framing','Norm of reciprocity', 'Low vs. high category scales') & outcomes1_2 == "mean _ SD") %>% 
  split(.$effect) %>% 
  map(~transform_SD(.$outcome_t1, .$outcome_c1, .$outcome_t2, .$outcome_c2, .$ntreatment, .$ncontrol)) %>% 
  map(~untruncated_d_to_r(.$d, .$n1, .$n2)) %>% 
  map_dfr(summarizer, .id = "effect")
```

```{r b-plot, cache = FALSE}
b_plot <- b_plot_1 + geom_point(data = res3, shape = 4, aes(x = I2, y = sqrt(tau2))) +
  scale_x_continuous(name = expression(paste(italic(I)^2, " Index"))) +
  scale_y_continuous(name = "Tau")

```

```{r tau-vs-tau}
d <- select(a, effect, tau2) %>% 
  left_join(select(b, effect, tau2), by = "effect")
```

```{r d-plot}
d_plot <- ggplot(d, aes(x = sqrt(tau2.x), y = sqrt(tau2.y))) + 
  geom_point() + 
  ggtitle("SMD vs. biserial") +
  scale_x_continuous(name = "Tau (SMD)") +
  scale_y_continuous(name = "Tau (biserial r)")

```

```{r, fig.width = 6, fig.asp = 1.3}
cowplot::plot_grid(a_plot, b_plot, d_plot, ncol = 1, labels = "AUTO")
```

_Figure S.1_ Association between $\tau$ and $I^2$ after and before standardized mean difference (SMD) effects were transformed into biserial correlations for all effects reported as SMDs in ManyLabs 1. Panel A shows the association between $\tau$ and $I^2$ without transformation when using the meta-analytic specification of the original authors (see main text). Panel B shows the same associations for the same effects after transformation to biserial correlations, either with truncation of correlations over 1 (circles) or without (crosses). Panel C shows the association between $\tau$ for the same effects before and after transformation.